Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2019-03-19 thesis/chapters/ch_introduction.tex
Updates to first half of introduction (to Stars section). Cleaned up text, added references, and included final bit in intro

%
% \beq
% \frac{\mathrm{d}}{\mathrm{d}t} \int_S \mathbf{B} \cdot \mathrm{d}S = 0,f
% \eeq

Galaxies are amalgamations of gas and stars embedded in dark matter halos, formed over cosmic time as the products of the hierarchical assembly that follows from the collapse of primordial density fluctuations that arose after the Big Bang. The gravitational pull of dark matter -- as predicted from $\Lambda$CDM cosmology -- controls the hierarchical growth of structure in the Universe. While, broadly speaking, the evolution of baryons is dominated by the gravitational pull of dark matter, the beautiful simplicity of $\Lambda$CDM cosmology is muddied by their existence. The complexities of hydrodynamics, magnetohydrodynamics, thermodynamics, chemistry, radiative processes, and nucleosynthesis -- or, ``astrophysics" for short -- drives these deviations and gives rise to the Universe that we observe and -- not least of all -- us. In spite of a concerted effort spanning nearly a century, we are far from a self-consistent theory of galaxy evolution \citep[see ][ for a recent review of outstanding problems in galactic evolution]{NaabOstriker2017}. It is understanding the rich set of physics that formed our own Galaxy and the countless galaxies scattered throughout the Universe that motivates this work.

A substantial slice of modern astrophysics has been devoted towards understanding the chemical evolution -- the abundances of individual elements in space and time -- of the Universe. With the exception of H, He, and trace amounts of light elements, all of the elements in the Universe are produced in nuclear reactions in the cores of stars, massive star stellar winds, asymptotic giant branch (AGB) winds of low mass stars, and more exotic sites like nuetron star - neutron star mergers (NS-NS). See \cite{Nomoto2013}, \cite{Thielmann2017}, and \cite{Frebel2018} for recent reviews on this topic as it applies to studies of galactic and stellar evolution. Much work has been done in studying the global metallicity evolution of galaxies through the mass-metallicity relationship \citep[e.g.][]{Lequeux1979,Tremonti2004,Lee2006,Zahid2012}, the metallicity gradient in our own and nearby galaxies \citep[e.g.][]{}, and detailed stellar abundances in galaxies with spectroscopically resolved stellar populations \citep[e.g.][]{}. Yet in spite of increased number and quality of observations of galactic chemical evolution, developing a complete model is still a daunting task.

%Developing a complete model for galactic chemical evolution is a daunting task.
To give an impression of the difficulties involved, here is a recipe for constructing a complete model of galactic chemical evolution. One must first perscribe a galaxy's connection to its large scale structure to understand the inflow of pristine, metal-free gas into the disk of the galaxy, and the accretion (via mergers) of stars formed previously in galaxies hosted by different dark matter halos. One can then worry about knowing where, how, and with what masses stars form in the first place from this gas, which requires understanding the hydrodynamic 
%and magnetohydrodynamic 
properties of the multi-scale, multi-phase turbulent interstellar medium (ISM). This in turn requires one to toss in a detailed understanding of the stellar feedback physics -- stellar radiation, stellar winds, and supernovae -- that helps to regulate star formation by destroying cold, star forming gas, driving the multi-phase structure of the ISM, and by driving outflows of gas into the circumgalactic medium (CGM) around galaxies. Simultaneously, one makes the simple step of nailing down a complete model of stellar structure, stellar evolution, the reaction rates and cross sections of every nuclear reaction, and the lifetimes of individual isotopes. Once all of this is understood, one can then color in the gas in the galaxy over time with the individual metal yields of stellar populations released over their lifetime. With a complete understanding of stellar feedback and the ISM, one can then be sure that they completely capture the mixing of these metals in the ISM over time, and will produce a stellar population with accurate metal abundance ratios. In addition, one would then produce realistic galactic winds, removing the correct amount of metals from the galaxy and enriching the CGM and the intergalactic medium (IGM). Finally, one can layer this model ad nauseum on top of a $\Lambda CDM$ model for the cosmological evolution of many galaxies, reproducing all observed properties of galactic chemical evolution as a function of both mass and redshift..... Wait... I forgot about magnetic fields and cosmic rays...

It should be obvious now to the reader why there remains so much to be learned about galactic chemical evolution. Developing a complete model for galactic chemical evolution is a daunting task. Yet it is clear that this field offers an incredibly exciting test of an incredible range of physics.

The observational landscape today is primed for developing a more detailed understanding these processes. Recent observational campaigns such as SEGUE \citep{Yanny2009}, RAVE \citep{Kunder2017}, the Gaia-ESO survey \citep{Gaia}, APOGEE and APOGEE2 \citep{APOGEE2010,APOGEE}, GALAH \citep{GALAH,Buder2018}, as well as upcoming observations, such as the Local Volume Mapper as part of SDSS-V, have generated tremendous amounts of information on detailed stellar abundnaces and stellar kinematics in our Milky Way and nearby Local Group dwarf galaxies. One of the most powerful proposed uses of this enormous trove of data is chemical tagging \citep{Freeman2002}, whereby stellar populations are analyzed in chemical space and 6D phase space to identify co-eval and co-natal groups of stars. This process of galactic archeaology aims to break down and identify each distinct stellar component of our Galaxy, explaining the process of their formation and evolution. Substantial work has been made recently to determine the efficacy of this approach \citep[e.g.][]{Armillotta2018}, yet the physical processes that give rise to stellar abundances as we observe them today are still uncertain.

@@

\textbf{need final par}

@@

We give a primer on how to model galactic evolution in hydrodynamics simulations. We focus on the physics needed to simulate galaxies (cooling, heating, chemistry, star formation, feedback, etc.) as it pertains to the aspects of galactic chemical evolution examined in this dissertation. This is meant to be a broad overview of the physical processes modelled in such simulations, focusing on aspects of their implementation rather than the derivation of theory behind such processes. While important, the latter is better reserved for a much more detailed work (i.e. a textbook), and focusing on the former gives a much more accurate represenation of the actual work done for this dissertation. This discussion includes a very brief overview of hydrodynamics and the numerical methods used in this work in Section~\ref{intro:sec:hydrodynamics}, nucleosynthesis and stellar evolution in Section~\ref{intro:sec:nucleosynthesis}, radiative cooling and chemistry (actual chemistry) in Section~\ref{intro:sec:cooling}, star formation and star particles in Section~\ref{intro:sec:sf} and Section~\ref{intro:sec:stars}, and stellar feedback in Section~\ref{intro:sec:feedback}.

But first, we note that Astronomy jargon is full of idiosyncrisies. As metioned before, the word ``chemical" in the ``galactic chemical evolution" is a bit of a misnomer. This is commonly applied simply to refer to studies interested in the evolution of metal abundances (either as a whole, or for individual isotopes) in galaxies over time, and less often to the astrochemistry that those elements may participate in. Likewise, ``metals" refers broadly to every element except H and He. Throughout this dissertation we often refer to the metallicity ($Z$) of a galaxy, its gas, or its stars. This quantity represents the total mass fraction of all metals -- and is computed as such in our simulations -- but we note that this is impossible to measure the abundances of all metals in astrophysical contexts outside our own Solar System. The metallicity of our own Sun ($Z_{\odot}$), for example, is still uncertain \citep{Asplund2009}. Instead, most observational works adopt Fe as a proxy for total metallicity in stars (due to its many, strong absorption lines) and O as a proxy in gas abundances (as it is the most abundant metal in the Universe and has convenient, strong emisson lines). These metallicity proxies are commonly reported in some form of normalized abundance ratio. For stellar metallicities, this is given in the form [A/B], where [A/B] = log($N_A / N_B$) - log($N_A / N_B$)$_{\odot}$, and $N$ refers to the number of atoms of a given element (usually as [Fe/H]). Confusingly, gas-phase abundances are commonly reported as just log($N_A / N_B$), often with the somewhat arbitrary normalization of 12, or log(O/H) + 12, for example.\footnote{And I am not going to mention issues in normalizatiion across methods of deriving these abundances... \citep[e.g.][]{KewleyEllison2008}.}

\subsection{Hydrodynamics}
Astrophysical hydrodynamics simulations are first distinguished by the numerical methods employed to solve the Euler equations that describe the time ($t$) evolution of the energy density ($E$), density ($\rho$), pressure ($p$), and peculiar velocity ($\bm{v}$) of a fluid:
\begin{align}
  \frac{\delta \rho}{\delta t} + \nabla \cdot \left(\rho \bm{v}\right)  &= 0,\\
  \frac{\delta \rho \bm{v}}{\delta t} + \nabla \cdot \left(\rho \bm{v}\bm{v} + \bm{I}p \right) &= 0,\\
  \frac{\delta E}{\delta t} + \nabla \cdot \left[\left(E +p\right)\bm{v}\right] &= 0.
\end{align}

Historically, codes fall into one of two camps: 1) Eulerian grid-based codes, including the popular codes / algorithms commonly in use (in some way) today such as \code{Zeus} \citep{StoneNorman1992}, \code{FLASH} \citep{FLASH}, \code{RAMSES} \citep{Teyssier2002}, \code{Athena} \citep{Athena}, \code{ART-II} \citep{Rudd2008}, and \code{Enzo} \citep{Enzo2014}, and 2) particle-based Lagrangian methods known as smooth particle hydrodynamics (SPH), such as \code{Gadget} \citep{Springel2005}, \code{PKDGRAV-2} \citep{Stadel2001}, \code{Gasoline} \citep{Wadsley2004}, and \code{Changa} \citep{Menon2015}. However, recent codes blur the lines between these distinctions with new algorithms for solving the fluid equations on a moving-mesh (e.g. \code{AREPO} \citep{Springel2010}), or meshless finite-mass / finite-volume methods (such as those implemented in \code{Gizmo} \citep{Hopkins2015}). Given the large variance in numerical methods, recent studies have examined the differences between these numerical implementations \citep[e.g.][]{Agertz2007}, including large-scale code comparison projects \citep[e.g.][]{AGORA,AGORA2}. While developing a notion of which code is ``correct" is an ill-posed problem, these studies have allowed for improvement across implementations and an insight into how numerical methods themselves drive uncertainty in our understanding of astrophysical problems.

In this Dissertation, we make use of the adaptive-mesh refinement (AMR) hydrodynamics code \code{Enzo}, as described in greater detail in Chapter~\ref{ch:chapter1}. As it pertains to galactic chemical evolution, however, the use of a grid-based code has the advantage that metals (which are advected as passive scalars that follow the fluid-flow in the simulations) are allowed to mix and diffuse naturally across fluid elements (grid cells). This is not possible in native SPH implementations, yet is necessary to reproduce realistic galactic chemical evolution properties as shown by multiple recent works testing implementations of diffusion in SPH simulations \citep[e.g.][]{Shen2010,Su2017a,Escala2018}. The disadvantage of the diffusion in \code{Enzo}, however, is that it is entirely numerical, and therefore its exact properties are challenging to characterize and are strongly resolution dependent. Throughout this dissertation we attempt to account for the effects of resolution on our results, but do not explicitly examine the properties of numerical diffusion itself.

@@

A variety of nucleosynthetic processes occur within stars and during their deaths that lead to the production of all of the elements of the periodic table. This is illustrated well in Figure~\ref{intro:fig:periodic table}, which shows the periodic table colored by the fractional importance of the nucleosynthetic processes that create each element. Due to the diversity in production sources across the periodic table, multi-element metal abundance measurements in stars can be used to to better understand and constrain how and where each metal is produced. Exactly which reactions occur where depend strongly on the initial abundances of elements in stars, their stellar masses, and, for heavy nuclei, the neutron density.

This means that elements from different nucleosynthetic channels are produced on different timescales during a galaxy's evolution, as set by the lifetimes of the stars in which they occur, and are released into the galaxy at different energies, with different spatial distributions within the galaxy, and at different rates. For example, O, an $\alpha$-element is produced predominately in massive stars and is therefore released via core collapse supernovae on timescales of $\sim$~10~Myr after a star formation event, near their stellar birth sites. By contrast, Sr is produced predominately via the s-process (discussed more below) which occurs in the asymptotic giant branch (AGB) winds of lower mass stars ($M < 8$~M$_{\odot}$) at much lower energies and on timescales of $\sim 100$~Myr - $1$~Gyr. By this time, these stars would have diffused farther away from their birth sites, potentially sampling a qualitatively different portion of the ISM than earlier core collapse SNe. Generally the differences in how different timescales affect observed stellar abundances patterns are well understood and easily modelled.\footnote{The canonical example of this is the ``knee" feature in a $[\alpha/$Fe] vs. [Fe/H] diagram, where the location of this knee signifies the point at which Type Ia enrichment (which produces significant Fe, but little to no $\alpha$ elements) begins to have a significant contribution over core collapse SNe (rich in $\alpha$ elements, but not Fe) \citep[e.g.[]{MatteucciBrocato1990, Geisler2005, Hill2018}. The location of this knee shifts in a galaxy depending on its SFH.} However, it is not well understood how differences in where and with what energies these events occur affect observed stellar abundance patterns. Investigating these differences is a key motivation of this dissertation.

As relevant to this work, we briefly summarize a few of the important nucleosynthetic channels shown in Figure~\ref{intro:fig:periodic table} and their sites. $\alpha$-elements, notably O, Mg, Si, Ca, and Ti, are produced during the lifetime of massive stars (O, Mg) or during the core collapse supernova explosion itself. In either case, these elements are synthesized through the capture of an $\alpha$ nucleus, progressing towards increasingly heavier nuclei in intervals of 4/2 in mass/atomic number (as $\alpha$ = $^{4}_2$He) from $^8_4$Be to the stable $^{52}_{26}$Fe and unstable $^{56}_{28}$Ni (which days to $^{56}_{26}$Fe). The iron group elements, notably Fe, Mn, Co, and Ni, are produced predominately in Type Ia supernovae, with variations depending on a single vs. double degenerate scenario and the initial composition of exploding white dwarf(s) (WD).

The presence of these heavier elements in stars enables the formation of even heavier nuclei through neutron capture. This is categorized into two processes, slow (or s-) and rapid (or r-) process. In the former, neutron densities are typically low enough that the product of each neutron capture beta decays to a more stable isotope before capturing another neucleon. In r-process, however, the rate of neutron capture exceeds beta decay, alowing for the production of different, typically heavier, nuclei. As shown in the diagram, ligher elements are produced via the s-process and heavier via the r-process, with notable overlap in certain elements, such as Ba.\footnote{For the sake of this discussion, the light neutron-capture primary process elemnts (sometimes referred to as ``lighter element primary process") should be considered s-process elements. This different nucleosynthetic channel, first identified in \cite{Travaglio2004}, arose from an apparent disagreement between models of s-process enrichment in galactic chemical evolution studies andnd abundances of elements including Sr, Y, and Zr. However there is still uncertainty as to whether or not these differences can be explained with different models of weak s-process (which can occur in the winds of massive stars) \citep[e.g.][]{Prantzos2018}.} The s-process occurs predominately in the low-energy winds of low mass (1-8 M$_{\odot}$) AGB stars on time scales of 100's of Myr up to 1 Gyr. Exactly which elements are produced in a given AGB star varies with stellar mass and depends strongly on metallicity, which determines the available heavy seed nuclei. Sr, Y, Zr, and Ba, as well as C, N and F, are commonly used as tracers of this nucleosynthetic channel.

The dominant origin of r-process enrichment is highly uncertain, though possible channels include neutron star - neutron star and neutron star - black hole mergers, netrino driven winds in core collapse SNe, and exotic SNe, such as magnetic, jet-driven SNe, collapsars, hypernovae, and long-duration gamma-ray bursts \citep[see ][ for recent reviews]{Frebel2018, Cowan2019}. Eu is the most commonly used (and easy to observe) unambigious observational tracer of r-process enrichment; Ba is also often studied as a tracer, but requires care due to the significant contribution from s-process enricment. More detailed spectra \citep[e.g.][]{Ji2018a} can trace many more of these elements, which can be used as an important discriminator between nucleosynthetic models.

Finally, although the production of elements unaffected by the r-process are comparatively well understood, there are still large uncertainties in both nuclear reaction rates and stellar evolution properties that affect the computed yields for any given element. As including self-consistent nuclear reactions in the interior of stars is nearly computationally impossible in galaxy-scale simulations, metal yields are generally included as averaged over an entire stellar population (see Section\ref{intro:sec:stars}) as determined from a tabulated set of yields. These uncertainties, however, mean there are often significant variations and inconsistencies between tables produced by separate groups, which can make some times make interpreting the results of galactic chemical evolution models challenging.

@@

In this dissertation, we develop the first implementation that breaks the SSP formalism entirely by following stars as individual star particles sampled from an adopted IMF. This allows us to place unprecidented detail on both the stellar feedback and metal yields associated with each star, spatially and temporally resolving differences in both of these channels as a function of stellar mass and metallicity.

@@

\subsection{Known Dwarf Galaxies of the Local Group}

@@

This dissertation utilizes high-resolution hydrodynamics simulations that are expensive to run in both walltime ($\sim$ weeks to months) and computational time ($\sim 10^{5-6}$ CPU hours). For this reason, it is computationally infeasible to directly explore many of the uncertainties associated with each of the components of the model as discussed in Section~\ref{intro:sec:ingredients}. For this reason, substantially cheaper semi-analytic models are valuable tools that can be used to better understand galactic evolution.

>>>>

%
% \beq
% \frac{\mathrm{d}}{\mathrm{d}t} \int_S \mathbf{B} \cdot \mathrm{d}S = 0,f
% \eeq
Galaxies are amalgamations of gas and stars embedded in dark matter halos. They are formed over cosmic time as the products of the hierarchical assembly that follows from the collapse of the primordial density fluctuations that arose after the Big Bang. The gravitational pull of dark matter -- as predicted from $\Lambda$CDM cosmology -- controls the hierarchical growth of structure in the Universe. While, broadly speaking, the evolution of baryons is dominated by this pull, the beautiful simplicity of $\Lambda$CDM cosmology is muddied by their existence. The complexities of hydrodynamics, magnetohydrodynamics, thermodynamics, chemistry, radiative processes, and nucleosynthesis -- or, ``astrophysics" for short -- drives these deviations and gives rise to the Universe that we observe. In spite of a concerted effort spanning nearly a century, we are far from a self-consistent theory of galaxy evolution (see \citet{SomervilleDave2015} and \cite{NaabOstriker2017} for recent reviews). It is understanding the rich set of physics that formed our own Galaxy and the countless galaxies scattered throughout the Universe that motivates this work.

A substantial slice of modern astrophysics has been devoted towards understanding the chemical evolution -- the abundances of individual elements in space and time -- of the Universe. With the exception of H, He, and trace amounts of light elements, all of the elements in the Universe are produced in nuclear reactions associated with stellar evolution, either in the cores of stars, stellar winds, asymptotic giant branch (AGB) winds of low mass stars, and more exotic sites like neutron star - neutron star mergers (NS-NS). See \cite{Nomoto2013}, \cite{Thielmann2017}, and \cite{Frebel2018} for recent reviews on this topic as it applies to studies of galactic and stellar evolution. Much work has been done in studying the global metallicity evolution of galaxies through the mass-metallicity relationship \citep[e.g.][]{Lequeux1979,Tremonti2004,Lee2006,Zahid2012,AndrewsMartini2013}, the metallicity gradient in our own and nearby galaxies \citep[e.g.][]{Searle1971,Shaver1983,Belfiore2017,SanchezMenguiano2017}, and detailed stellar abundances in nearby dwarf galaxies with spectroscopically resolved stars (see the review in \cite{Tolstoy2009}). Yet in spite of increased number and quality of observations of galactic chemical evolution, developing a complete model of galactic chemical evolution is still a daunting task.

%Developing a complete model for galactic chemical evolution is a daunting task.
To give an impression of the difficulties involved, we outline a recipe for constructing a complete model of galactic chemical evolution. One must first prescribe a galaxy's connection to its large scale structure to understand the inflow of pristine, metal-free gas into the disk of the galaxy, and the accretion (via mergers) of stars formed previously in galaxies hosted by different dark matter halos. One can then worry about knowing where, when, how, and with what masses stars form in the first place from this gas. This requires understanding the hydrodynamic
%and magnetohydrodynamic
properties of the multi-scale, multi-phase turbulent interstellar medium (ISM) and the radiative processes and chemistry that operate within the ISM. This in turn requires one to toss in a detailed understanding of the stellar feedback physics -- stellar radiation, stellar winds, and supernovae -- that helps to regulate star formation by destroying cold, star forming gas, driving the multi-phase structure of the ISM, and by driving outflows of gas into the circumgalactic medium (CGM) around galaxies. Simultaneously, one makes the simple step of nailing down a complete model of stellar structure, stellar evolution, the reaction rates and cross sections of every nuclear reaction, and the lifetimes of individual isotopes. Once all of this is understood, one can then color in the gas in the galaxy over time with the individual metal yields of stellar populations released over their lifetime. With a complete understanding of stellar feedback and the ISM, one can then be sure that they completely capture the mixing of these metals in the ISM over time, and will produce a stellar population with accurate metal abundance ratios. In addition, one would then produce realistic galactic winds, removing the correct amount of metals from the galaxy and enriching the CGM and the intergalactic medium (IGM). Finally, one can layer this model ad nauseum on top of a $\Lambda CDM$ model for the cosmological evolution of many galaxies, reproducing all observed properties of galactic chemical evolution as a function of both mass and redshift..... Wait... I forgot about magnetic fields, active galactic nuclei, and cosmic rays...

It should be obvious now to the reader why there remains so much to be learned about galactic chemical evolution. Yet it is clear that this field offers an incredibly exciting test of a wide range of physical processes.

The observational landscape today is primed for developing a more detailed understanding these processes. Recent observational campaigns such as SEGUE \citep{Yanny2009}, RAVE \citep{Kunder2017}, the Gaia-ESO survey \citep{Gaia}, APOGEE and APOGEE2 \citep{APOGEE2010,APOGEE}, GALAH \citep{GALAH,Buder2018}, as well as upcoming observations, such as the Local Volume Mapper as part of SDSS-V, have generated tremendous amounts of information on detailed stellar abundances and stellar kinematics in our Milky Way and nearby Local Group dwarf galaxies. One of the most powerful proposed uses of this enormous trove of data is chemical tagging \citep{Freeman2002}, whereby stellar populations are analyzed in chemical space and 6D phase space to identify co-eval and co-natal groups of stars. This process of galactic archeology aims to break down and identify each distinct stellar component of our Galaxy, explaining the process of their formation and evolution. Substantial work has been made recently to determine the efficacy of this approach \citep[e.g.][]{Ting2012,Hogg2016,Jofre107,Price-Jones2018,Armillotta2018}, yet the physical processes that give rise to stellar abundances as we observe them today are still uncertain.

% I could maybe add a paragraph here mentioning work on galactic outflows and metals in the CGM and how studying
% how this occurs is very important (peeples metal census?)

@@

In order to better understand galactic chemical evolution as a whole, we study the detailed process of metal enrichment in the ISM and the ejection of metals into the CGM through high-resolution hydrodynamics simulations of individual galaxies. In the remainder of this introduction, we discuss the physical process that we consider in constructing a complete model of galactic evolution in Section~\ref{intro:sec:ingredients} with an emphasis on how these processes are treated in simulations, a summary of relevant observations of the nearby dwarf galaxies relevant to this study in Section~\ref{intro:sec:dwarf galaxies}, a discussion of alternative approaches to modeling galactic chemical evolution in Section~\ref{intro:sec:onezone}, and an outline of the rest of this \dissertation in Section~\ref{intro:sec:structure}.

@@

The following is a primer on how to model galactic evolution in hydrodynamics simulations. We focus on the physics needed to simulate galaxies (cooling, heating, chemistry, star formation, feedback, etc.) as it pertains to the aspects of galactic chemical evolution examined in this \dissertation. This is meant to be a broad overview of the physical processes treated in such simulations, focusing more on aspects of their implementation rather than the derivation of theory behind each process. While important, the latter is beyond the scope of this introduction and would readily turn into a full textbook. Focusing on implementation gives a much more accurate representation of the actual work done and skills learned during work on this \dissertation. This discussion includes a very brief overview of hydrodynamics and the numerical methods used in this work in Section~\ref{intro:sec:hydrodynamics}, nucleosynthesis and stellar evolution in Section~\ref{intro:sec:nucleosynthesis}, radiative cooling and chemistry (actual chemistry) in Section~\ref{intro:sec:cooling}, star formation and star particles in Section~\ref{intro:sec:sf} and Section~\ref{intro:sec:stars}, and stellar feedback in Section~\ref{intro:sec:feedback}.

But first, we note that Astronomy jargon is full of idiosyncrasies. As mentioned before, the word ``chemical" in ``galactic chemical evolution" is a bit of a misnomer. This is commonly applied simply to refer to studies interested in the evolution of metal abundances (either as a whole, or for individual isotopes) in galaxies over time, and less often to the chemical reactions that those elements may participate in. Unfortunately ``chemistry" is often used with both meanings in the same work requiring context to decipher the intended meaning. Likewise, ``metals" refers broadly to every element except H and He. Throughout this \dissertation we often refer to the metallicity ($Z$) of a galaxy, its gas, or its stars. This quantity represents the total mass fraction of all metals -- and is computed as such in our simulations -- but we note that it is generally impossible to measure the abundances of all metals in astrophysical contexts outside our own Solar System (and even then, it is challenging). The metallicity of our own Sun ($Z_{\odot}$), for example, is still uncertain \citep{Asplund2009}. Instead, most observational works adopt Fe as a proxy for total metallicity in stars (due to its many, strong absorption lines), O as a proxy in gas-phase abundances in the ISM (as it is the most abundant metal in the Universe and has convenient, strong emission lines), and O or C in gas-phase abundances of the CGM when observed in absorption (primarily through the lines OVI and CIV). These metallicity proxies are usually reported in some form of normalized abundance ratio. For stellar metallicities, this is given in the form [A/B]. [A/B] = log($N_A / N_B$) - log($N_A / N_B$)$_{\odot}$, where $N$ refers to the number of atoms of a given element (e.g. [Fe/H]). Confusingly, gas-phase abundances are commonly reported as just log($N_A / N_B$), often with the somewhat arbitrary normalization of +12, or log(O/H) + 12, for example; converting between the two definitions is straightforward (albeit annoying) and requires adopting a value for the solar abundance. \footnote{This is not to mention issues in normalization across methods of deriving these abundances... \citep[e.g.][]{KewleyEllison2008}.}

\subsection{Hydrodynamics}
\label{intro:sec:hydrodynamics}

Astrophysical hydrodynamics simulations are first differentiated by the numerical methods they employ to solve the Euler equations. These equations describe the time ($t$) evolution of the energy density ($E$), density ($\rho$), pressure ($p$), and peculiar velocity ($\bm{v}$) of a fluid:
\begin{align}
  \frac{\delta \rho}{\delta t} + \nabla \cdot \left(\rho \bm{v}\right)  &= 0,\\
  \frac{\delta \rho \bm{v}}{\delta t} + \nabla \cdot \left(\rho \bm{v}\bm{v} + \bm{I}p \right) &= 0,\\
  \frac{\delta E}{\delta t} + \nabla \cdot \left[\left(E +p\right)\bm{v}\right] &= 0.
\end{align}

Historically, codes fall into one of two camps: 1) Eulerian grid-based codes, including the popular codes / algorithms commonly in use (in some way) today such as \code{Zeus} \citep{StoneNorman1992}, \code{FLASH} \citep{FLASH}, \code{RAMSES} \citep{Teyssier2002}, \code{Athena} \citep{Athena}, \code{ART-II} \citep{Rudd2008}, and \code{Enzo} \citep{Enzo2014}, and 2) particle-based Lagrangian methods known as smooth particle hydrodynamics (SPH), such as \code{Gadget} \citep{Springel2005}, \code{PKDGRAV-2} \citep{Stadel2001}, \code{Gasoline} \citep{Wadsley2004}, and \code{Changa} \citep{Menon2015}. However, recent codes blur the lines between these distinctions with new algorithms for solving the fluid equations on a moving-mesh (e.g. \code{AREPO} \citep{Springel2010}), or meshless finite-mass / finite-volume methods (such as those implemented in \code{Gizmo} \citep{Hopkins2015}). Traditionally these codes were designed to run exclusively on CPUs, but significant work has been made recently to offload various portions onto GPUs (for example \code{GAMER} and \code{GAMER-2} \citep{Schive2010,Schive2018}, \code{GPUGAS} \code{Kulikov2014}, and \code{CHOLLA} \citep{Schneider2015}). Given the large variance in numerical methods, recent studies have examined the differences between these numerical implementations \citep[e.g.][]{Agertz2007}, including large-scale code comparison projects \citep[e.g.][]{AGORA,AGORA2}. While developing even a notion of which code is ``correct" is an ill-posed problem, these studies have allowed for improvement across implementations and an insight into how numerical methods themselves drive uncertainty in our understanding of astrophysical problems.

In this \dissertation, we make use of the adaptive-mesh refinement (AMR) hydrodynamics code \code{Enzo}, as described in greater detail in Chapter~\ref{ch:chapter1}. As it pertains to galactic chemical evolution, however, the use of a grid-based code has the advantage that metals (which are advected as passive scalars that follow the fluid-flow in the simulations) are allowed to mix and diffuse naturally across fluid elements (grid cells). This is not possible in native SPH implementations, yet is necessary to reproduce realistic galactic chemical evolution properties as shown by multiple recent works testing implementations of diffusion in SPH simulations \citep[e.g.][]{Shen2010,Su2017a,Escala2018}. The disadvantage of the diffusion in \code{Enzo}, however, is that it is entirely numerical, and therefore its exact properties are challenging to characterize and are strongly resolution dependent. Throughout this \dissertation we attempt to account for the effects of resolution on our results, but do not explicitly examine the properties of numerical diffusion itself.

@@

A variety of nucleosynthetic processes occur within stars and during their deaths that lead to the production of all of the elements in the periodic table. This is illustrated well in Figure~\ref{intro:fig:periodic table}, which shows the periodic table colored by the fractional importance of the nucleosynthetic processes that produce each element. Due to the diversity in production sources across the periodic table, multi-element metal abundance measurements in stars can be used to to better understand and constrain how and where each metal is produced. Exactly which reactions occur where depend strongly on the initial abundances of elements in stars, their stellar masses, and, for heavy nuclei, the neutron density.

This means that elements from different nucleosynthetic channels are produced on different timescales during a galaxy's evolution, as set by the lifetimes of the stars in which they occur, and are released into the galaxy at different energies, with different spatial distributions within the galaxy, and at different rates. For example, O, an $\alpha$-element is produced predominately in massive stars and is therefore released via core collapse supernovae on timescales of $\sim$~10~Myr after a star formation event, near their stellar birth sites. By contrast, Sr is produced predominately via the s-process (discussed more below) which occurs in part in the asymptotic giant branch (AGB) winds of lower mass stars ($M < 8$~M$_{\odot}$) at much lower energies and on timescales of $\sim 100$~Myr - $1$~Gyr. By this time, these stars would have diffused farther away from their birth sites, potentially sampling a qualitatively different portion of the ISM than the preceeding core collapse SNe. Generally, how different timescales affect observed stellar abundances patterns are well understood and easily modeled.\footnote{The canonical example of this is the ``knee" feature in a $[\alpha/$Fe] vs. [Fe/H] diagram, where the location of this knee signifies the point at which Type Ia enrichment (which produces significant Fe, but little to no $\alpha$ elements) begins to have a significant contribution over core collapse SNe (rich in $\alpha$ elements, but not Fe) \citep[e.g.[]{MatteucciBrocato1990, Geisler2005, Hill2018}. The location of this knee in this diagram shifts for a given galaxy depending on its SFH.} However, it is not well understood how differences in where and with what energies these events occur affect observed stellar abundance patterns. Investigating these differences is a key motivation of this \dissertation.

As relevant to this work, we briefly summarize a few of the important nucleosynthetic channels shown in Figure~\ref{intro:fig:periodic table} and their sites. $\alpha$-elements, notably O, Mg, Si, Ca, and Ti, are produced during the lifetime of massive stars (O, Mg) or during the core collapse supernova explosion itself. In either case, these elements are synthesized through the capture of an $\alpha$ nucleus, progressing towards increasingly heavier nuclei in intervals of 4/2 in mass/atomic number (as $\alpha$ = $^{4}_2$He) from $^8_4$Be all the way to the stable $^{52}_{26}$Fe and unstable $^{56}_{28}$Ni (which days to $^{56}_{26}$Fe). The iron group elements, notably Fe, Mn, Co, and Ni, are produced predominately in Type Ia supernovae, with variations depending on a single vs. double degenerate scenario, the initial composition of exploding white dwarf(s) (WD), and the dynamics of the explosion itself.

The presence of these heavier elements in stars enables the formation of even heavier nuclei through neutron capture. This is categorized into two processes, the slow (or s-) and rapid (or r-) process. In the former, neutron densities are typically low enough that the product of each neutron capture beta decays to a more stable isotope before capturing another nucleon. In r-process, however, the rate of neutron capture exceeds beta decay, allowing for the production of different, typically heavier, nuclei. As shown in the diagram, lighter elements are produced via the s-process and heavier via the r-process, with notable overlap in certain elements, such as Ba.\footnote{For the sake of this discussion, the light neutron-capture primary process elements (sometimes referred to as ``lighter element primary process") should be considered s-process elements. Briefly, this different nucleosynthetic channel, first identified in \cite{Travaglio2004}, arose from the disagreement between models of s-process enrichment and stellar abundances of elements including Sr, Y, and Zr. However there is still uncertainty as to whether or not these differences can be explained with different models of weak s-process (which can occur in the winds of massive stars) \citep[e.g.][]{Prantzos2018}.} The s-process occurs predominately in the low-energy winds of low mass (1-8 M$_{\odot}$) AGB stars on time scales of 100's of Myr up to 1 Gyr. Exactly which elements are produced in a given AGB star varies with stellar mass and depends strongly on metallicity, which determines the available heavy seed nuclei. Sr, Y, Zr, and Ba, as well as C, N and F, are commonly used as observational tracers of this nucleosynthetic channel.

The dominant origin of r-process enrichment is highly uncertain, though possible channels include NS-NS and neutron star - black hole mergers, neutrino driven winds in core collapse SNe, and exotic SNe, such as magnetic, jet-driven SNe, collapsars, hypernovae, and long-duration gamma-ray bursts \citep[see ][ for recent reviews]{Frebel2018, Cowan2019}. Eu is the most commonly used (and easy to observe) unambiguous observational tracer of r-process enrichment; Ba is also often studied as a tracer, but requires care due to the significant contribution from s-process enrichment. More detailed spectra \citep[e.g.][]{Ji2018a} can trace many more of these elements, which can be used as an important discriminator between nucleosynthetic models.

Finally, although the production of elements unaffected by the r-process are comparatively well understood, there are still large uncertainties in both nuclear reaction rates and stellar evolution properties that affect the computed yields for any given element. As including self-consistent nuclear reactions in the interior of stars is nearly computationally impossible in galaxy-scale simulations, metal yields are generally included as averaged over an entire stellar population (see Section\ref{intro:sec:stars}) as obtained from a tabulated set of yields. These yields for a given star in a simulation are typically IMF-averaged, and occasionally mass and metallicity dependent depending on the level of detail desired in the simulation. Often, only the global metallicity is tracked in simulations, but tracking individual metal species (especially C, O, and Si which can affect the chemistry and cooling physics in the ISM) is becoming more common. The uncertainties across tabulated yields, however, mean there are often significant variations and inconsistencies between tables produced by separate groups. These can sometimes make interpreting the results of galactic chemical evolution models challenging.

@@

In this \dissertation, we develop the first implementation that breaks the SSP formalism entirely by following stars as individual star particles sampled from an adopted IMF. This allows us to place unprecidented detail on both the stellar feedback and metal yields associated with each star, spatially and temporally resolving differences in both of these channels as a function of stellar mass and metallicity.

@@

\subsection{Known Dwarf Galaxies of the Local Group}
\label{intro:sec:dwarf galaxies}

@@

This \dissertation utilizes high-resolution hydrodynamics simulations that are expensive to run in both walltime ($\sim$ weeks to months) and computational time ($\sim 10^{5-6}$ CPU hours). For this reason, it is computationally infeasible to directly explore many of the uncertainties associated with each of the components of the model as discussed in Section~\ref{intro:sec:ingredients}. For this reason, substantially cheaper semi-analytic models are valuable tools that can be used to better understand galactic evolution.
